---
title: "PA1_template"
author: "ande4148"
date: "December 20, 2015"
output: md_document
---

Load libraries:

```{r, results='hide'}
library("ggplot2", lib.loc="C:/Users/aeand/R/R-3.2.2/library")
library("dplyr", lib.loc="C:/Users/aeand/R/R-3.2.2/library")
library("lattice", lib.loc="C:/Users/aeand/R/R-3.2.2/library")
```

Download dataset:

```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(fileUrl,destfile="repdata_data_activity.zip")
unzip(zipfile="repdata_data_activity.zip")
```

Load data:

```{r}
repdata <- read.csv("activity.csv")
```

Create data frame of the total number of steps taken each day:

- Manipulate date and interval elements 
- Sum the intervals for each day
- Create and label data frame with dates and steps

```{r}
repdata$date <- as.Date(as.character(repdata$date))
repdata$interval <- as.factor(repdata$interval)
daysteps <- tapply(repdata$steps, factor(repdata$date), sum)
x <- as.factor(unique(repdata$date))
daysum <- data.frame(x, daysteps)
names(daysum) <- c("Date", "Total Steps")
```

Create histogram of total steps taken each day:

```{r}
plot(daysum$Date, daysum$`Total Steps`)
lines(daysum$Date, daysum$`Total Steps`, type = "h")
```

Calcluate mean and median of total steps take each day:

```{r}
meansteps <- mean(daysteps, na.rm = T)
medsteps <- median(daysteps, na.rm = T)
meansteps
medsteps
```

Calculate the mean steps taken for each interval accross all days:

- Calculate means
- Create data frame

```{r}
meanalldays <- data.frame(tapply(repdata$steps, factor(repdata$interval), mean, 
                                na.rm = T))
names(meanalldays) <- c("Mean.All.Days")
```

Create time series plot for mean steps taken by interval:

```{r}
labels <- c("2am", "4am", "6am", "8am", "10am", "12pm", "2pm", "4pm", "6pm", "8pm", 
            "10pm", "12am")
at <- seq(24, 288, by = 24)
plot(meanalldays, xlab = "Interval", ylab = "Mean Steps Across All Days", xrange 
     = meanalldays$Interval, type = "l", xaxt = "n")
axis(side = 1, at = at, labels = labels)
```

Find and print the interval and value of the interval with the most mean steps taken

```{r}
max <- max(meanalldays)
meanalldays[meanalldays$Mean.All.Days == max,]
```

Impute missing values in dataset (all missing values are in 'steps' column):

- Sum missing values
- Subset original data rows with missing values
- Create data frame of mean steps taken for each interval
- Replace missing values with the mean steps taken in corresponding interval
- Combine data frames into new data set with missing values imputed

```{r}
sum(is.na(repdata))
nasub <- subset(repdata, is.na(repdata$steps))
nonnasub <- subset(repdata, !is.na(repdata$steps))
intermeans <- meanalldays
intermeans$interval <- levels(repdata$interval)
nasub$steps[nasub$interval == intermeans$interval] <- intermeans$Mean.All.Days
repdata.i <- rbind(nonnasub, nasub)
```

Recalculate total steps taken for each day using imputed data set:

- Sum the intervals for each day
- Create and label data frame with dates and steps

```{r}
daysteps.i <- tapply(repdata.i$steps, factor(repdata.i$date), sum)
x.i <- as.factor(unique(repdata.i$date))
daysum.i <- data.frame(x.i, daysteps.i)
names(daysum.i) <- c("Date", "Total Steps")
```

Create histogram of total steps taken each day, with missing values imputed:

```{r}
plot(daysum.i$Date, daysum.i$`Total Steps`)
lines(daysum.i$Date, daysum.i$`Total Steps`, type = "h")
```

Calcluate mean and median of total steps take each day, with missing values imputed:

```{r}
meansteps.i <- mean(daysteps.i)
medsteps.i <- median(daysteps.i)
meansteps.i
medsteps.i
```

Compare difference in means and medians for original and imputed datasets:

- Print the values for the imputed data set minus the values for the origianl dataset

```{r}
meansteps.i - meansteps
medsteps.i - medsteps
```

Create separate data frames for weekday and weekend observations,using dataset with imputed values:

- Label observations by day of the week
- Split into data frame for weekdays and data frame for weekend days
- Create day type variable for weekdays and weekend days

```{r}
repdata.i$wkdays <- weekdays(repdata.i$date, abbreviate = T)
d <- c("Mon", "Tue", "Wed", "Thu", "Fri")
wkdaydf <- subset(repdata.i, repdata.i$wkdays %in% d)
wkenddf <- subset(repdata.i, !repdata.i$wkdays %in% d)
wkdaydf$daytype <- "weekday"
wkenddf$daytype <- "weekend"
```

Calculate the mean steps taken for each interval across days for day type (weekday and weekend):

- Calculate means steps for intervals across weekday day type
- Create data frame with means steps for each interval 

```{r}
meanwkdays <- data.frame(tapply(wkdaydf$steps, factor(wkdaydf$interval), mean))
names(meanwkdays) <- "Mean.Wk.Days"
meanwkdays$interval <- levels(repdata$interval)
wkdaydf$steps[wkdaydf$interval == meanwkdays$interval] <- meanwkdays$Mean.Wk.Days
```

- Calculate means for wintervals across weekend day type
- Create data frame with means steps for each interval 

```{r}
meanwkenddays <- data.frame(tapply(wkenddf$steps, factor(wkenddf$interval), mean))
names(meanwkenddays) <- "Mean.Wk.Days"
meanwkenddays$interval <- levels(repdata$interval)
wkenddf$steps[wkenddf$interval == meanwkenddays$interval] <- meanwkenddays$Mean.Wk.Days
```

- Combine weekday and weekend data frames into new data set
- Convert day type labels to factors

```{r}
repdata.i2 <- rbind(wkdaydf, wkenddf)
repdata.i2$daytype <- as.factor(repdata.i2$daytype)
```

Create panel of time series plots for mean steps taken by interval for each day type:

```{r}
xyplot(steps ~ interval | daytype, data = repdata.i2, type = "l", 
       scales = list(x=list(at = at, labels = labels)))
```





















